How One Microtask Affects Another

### 2015-09-22 docs/publications/2015.08_CHI/man.tex
chi manuscript nearly complete

% My packages
\usepackage{dcolumn}
\usepackage{multirow}
\newcolumntype{d}{D{.}{.}{4.0}}
\newcolumntype{s}{D{.}{.}{1.4}}
\usepackage{arydshln}

@@

\begin{abstract}
Microtask platforms are becoming commonplace tools for performing human
research, producing gold-standard data, and annotating large datasets.
These platforms connect \textit{requesters}
(researchers or companies) with large populations (crowds) of workers, who 
perform small jobs, typically taking less than five minutes each.
A topic of ongoing research concerns designing jobs that elicit high
quality annotation.
Here we identify a feature of nearly all crowdsourcing 
jobs that profoundly impacts workers' responses.
Microtask jobs typically consist of a sequence 
of tasks sharing common format 
(e.g., circle galaxies in an image). 
Using image-labeling, a canonical microtask job format, we 
discover that earlier tasks have a priming effect on the worker, shifting 
the distribution of future responses by 30-50\% 
(total variational distance). 
Moreover, prior tasks influence the content that workers focus on, 
as well as the richness and specialization of responses. 
In comparison to overtly framing the job, by stating the requester's 
research interest, we find prior-task effects are on par or stronger
than those of framing.
While intertask effects can be a source of systematic bias, 
our results suggest that, with appropriate task design, 
they might be leveraged to hone worker focus and acuity, 
helping to elicit reproducible, expert-level judgments.
Prior-task effects are a crucial aspect of human computation that should be
considered in the design of any crowdsourced study.
\end{abstract}

@@

\section{Introduction}
There are many tasks that are trivial for people, but difficult to solve
programatically.  
Typical tasks include tagging and categorizing images
%\cite{6116320,Zhai2012357}, 
coding and transcribing media,
%\cite{chandler2013breaking,Berinsky2012351,Finnerty2013},%paolacci2010running},
judging the relevancy or quality of content, and performing surveys for 
academic or market reserach purposes\footnote{We provide a survey of task 
types in the supplementary text.}.
%\cite{le2010ensuring,grady2010crowdsourcing,alonso2009can,kazai2013analysis}.
Often these tasks are ill-defined and require some kind of high-level 
judgment.
Microtask platforms are marketplaces that help fill the gap in current 
computational capabilities, by matching requesters, who need to have such 
tasks completed, with human workers.  Amazon's Mechanical Turk, and 
CrowdFlower are two popular microtask platforms.

This form of crowdsourcing embeds workers in a controlled but flexible
task infrastructure.  To the requester, workers seem almost like input-output 
devices.  
This provides much of the flexibility and 
cost-savings of fully automating the work in a computer program: 
the workforce is available on demand over the Internet using automated 
scripts, without the need for interviews or contracts 
\cite{wolfson2011look,5543192}.
Jobs can be performed at a fraction of the cost of traditional methods for 
recruiting temporary workers or experimental subjects
\cite{Berinsky2012351}. %,ranade2012crowdsource}%,paolacci2010running}.
Many researchers consider microtask platforms 
as a new kind of \textit{human computing} architecture
\cite{5543192}. %,little2010turkit,minder2012crowdlang,kittur2011crowdforge}.

The flexibility and cost-effectiveness of microtask labour
has led to a surge in demand from industry and 
academia\cite{wolfson2011look,Berinsky2012351}.  More recently microtask 
platforms have
been assessed as a way to supplement expert human resources to increase
capacity in critical applications, such as in medical diagnostic functions,
with promising results \cite{Warby2014385}.
Naturally, researchers have investigated the factors affecting the reliability
of microtask work, including
the level of pay \cite{Mason200977,kazai2013analysis}, 
the training and pre-screening of workers 
\cite{le2010ensuring,kazai2013analysis}, %paolacci2010running}, 
user-interface design \cite{Finnerty2013}, and the provision of timely 
feedback \cite{Dow20121013}.
Researchers have also investigated \textit{framing}, 
by altering the description of the workflow context 
\cite{Kinnaird2012281}, the stated purpose of tasks 
\cite{chandler2013breaking}, and the problem description
\cite{thibodeau2013natural}.  Another important optimization concerns
workflow design: how breaking down a complex problem into microtasks
\cite{kittur2011crowdforge}, altering microtask size 
\cite{Huang201077}, and introducing workflow interruptions 
\cite{laseckieffects} impact worker quality and efficiency.

Here, we focus on a ubiquitous feature of microtask work: the tendency
for workers to perform many similar tasks in quick succession.  
This tendency arises from a combination of worker preferences and the 
logistics of serving small tasks.  
Workers have a preference for performing sequences of similar microtasks
\cite{Chilton20101}, probably because it reduces 
cognitive load from task-switching\cite{Adamczyk2004271}.
Morever, as workers complete task assignments, they must continually 
switch between working on an assignment and choosing their next 
assignment (weighing such factors as the wage paid, and effort required).  
Bundling tasks together reduces the amount of time spent choosing 
assignments and reduces the overhead of switching between working and 
choosing.  This may explain why the the pre-defined templates offered to 
help requesters set up their microtask jobs generally bundle many tasks 
together by default\footnote{mturk.com}.

The effects of task-repetition have been studied in the context of 
monitoring tasks by facility operators.  There the main concern is 
operator error due to boredom and distraction.  
However, in the microtask context, tasks are often ill-defined, and lack 
a clear notion of error.  For example for an image-labeling task
it would be difficult to decide which labels are `correct'.
Not only does this complicate the measurement of task-repitition effects, 
it suggests that their effect may be better understood as introducing 
\textit{bias} into responses, rather than \textit{error}.

One of our present contributions is a method for determining
the extent of this bias.  Generally, there may be many possible responses 
to an ill-defined, qualitative microtask, so the responses actually
obtained can be viewed as being drawn from a distribution.
Design factors that affect how workers respond to a microtask will change 
this distribution.  Measuring the effect of such a design factor therefore
amounts to measuring how much the distribution of responses changes
There are good statistical techniques for detecting \textit{whether}
samples come from two different distributions, such as the $\chi^2$ test.  
But to properly weigh design decisions, a method is needed to measure
the \textit{extent} by which response distributions differ.
The direct substitution of sample frequencies in divergence 
calculations, although occaisionally seen in literature, has no statistical
basis and can easily lead to the ``detection'' of large changes when none 
exist.  We present a convenient and statistically sound method that can be 
applied even when the set of possible responses is very large 
(or infinite), such as for free text responses.  
While our purpose is to measure the 
effects of prior microtasks on later ones, this method can be used to 
calculate an empirical divergence of distributions from samples in any 
context.

[\#\#]
Unlike tasks that have a clear correct response, tasks that rely on 
worker's qualitative judgments may be particularly susceptible to 
\textit{priming effects}.
Priming is the phenomenon by which a person's response to a prompt is
influenced by their recent exposure to stimuli
\cite{BJOP1796}. %,No2007}. %,beller1971priming}.  
Priming can arise when 
a prior stimulus is conceptually or perceptually related to an ensuing 
task \cite{Gass1999549} %sohn2001task}
so the repetition of similar tasks creates many opportunities for 
strong priming effects. 
If tasks can act as primes for other tasks, it may help explain an earlier 
observation [*] that when interrupting one microtask with another,
the interruption is more disruptive when the tasks are \textit{similar} to 
one another.

%Researchers investigating a somewhat similar class of psychological phenomena,
%found that ratings that microtask workers give to items in a list
%depend on the order in which the list is presented, an effect known as
%\textit{positional bias} \cite{lerman2014leveraging}.  This serves as an 
%example of a basic psychological phenomenon impacting crowdsourced data,
%and raises questions about what other cognitive biases might
%affect microtask work, and how prevalent the induced bias might be.

We investigate the potential for task repetition to induce bias in 
a highly generalizeable microtask setup, in which workers label a 
series of images.  If task-induced priming arises in our setup, 
it should be expected in virtually every microtask and crowdsourced workflow
where workers are called upon to exercise their judgment\footnote{
	As opposed to those tasks where a single clear correct answer exists.
}.
Our results show that a worker's response during a task
does indeed depend strongly on the tasks performed beforehand.   
We call the effects that earlier tasks exert on later ones 
\textit{intertask effects}.  If, as has been suggested \cite{5543192}, 
microtask platforms are to be considered as a new form of computing 
architecture, it will be necessary to reconcile the fact that the human 
computing elements 
exhibit \textit{hysteresis}, meaning that microtask workers' outputs depend 
on the \textit{history} of their inputs.

%As we have described, microtask platforms have been widely adopted as a 
%tool, but computer scientists also hold such platforms as the object of 
%research, viewing  microtask platforms as a new kind of computing 
%architecture.  In analogy to CPUs (central 
%processing units) and GPUs (graphics processing units), 
%Davis \textit{et al} invoked the \textit{human processing unit}, or HPU \cite{5543192}.
%Researchers have sought to define a set of basic HPU operations, and
%have developed libraries and frameworks to simplify HPU programming 
%and make it more like CPU programming 
%\cite{little2010turkit,minder2012crowdlang,kittur2011crowdforge}.  
%As an example, one operation overcomes the inherent stochasticity of HPU 
%outputs, by automatically aggregating votes from multiple HPUs under a single
%function call.  
%
%But our revelation of strong intertask effects shows that, 
%in addition to stochasticity, HPUs are subject to \textit{hysteresis}, 
%meaning that their 
%outputs depend on the history of their inputs.  Reproducibility and 
%path-independence are crucial properties for the components of a 
%computing architecture, so a better understanding of the computational 
%ramifications of many psychological phenomena is needed.

Our image-labeling tasks were performed on the MTurk microtask platform.  
We chose image labeling because it is very common\footnote{
	We discuss task composition in the supplementary text.
},
%\cite{chandler2013breaking,Finnerty2013},%,paolacci2010running},
and because it is prototypical for any kind of task in which the worker is 
asked to provide a response that is not precisely determined by 
the prompt, but instead depends on the worker's judgment.  
This includes such diverse tasks as describing video and audio recordings, 
summarizing texts, or in clinical diagnostic judgments.

In our first experiment, called \textit{intertask-food-objects},
workers were assigned to either a
\textit{food} or \textit{objects} treatment.  Workers in both treatments
performed a set of five \textit{initial tasks}, which each involved 
providing descriptive labels for an image depicting 
either food or (non-food) objects, depending on the worker's 
treatment (Fig.~\ref{fig:task}A and B).  Following the initial 
tasks, workers performed five \textit{test tasks}, which were identical for 
both treatments.  In the the test tasks, workers labeled images, each of 
which depicted a mixture of food and objects (Fig.~\ref{fig:task}C).  
The five initial 
and test tasks performed by each worker comprised one HIT\footnote{
	Workers were only allowed to participate in one HIT relating to this study.
}. The images were presented one at a time, and we made no 
distinction between the initial and test tasks.

The frequencies of words used by
workers in the \textit{food} and \textit{objects} treatments differed
significantly during test tasks (by $\chi^2$ test, $p = 6.0 \times 10^{-14}$),
even though the test tasks were identical in both treatments\footnote{
	see supplementary text for tabulated statistics.
}.  Thus earlier tasks do significantly affect later ones, raising concerns 
about the the microtask methodology.

Having affirmed the \textit{existence} of intertask effects, we wish to
understand the nature and severity of the bias induced.
We therefore define the extent of bias, $\theta$, in terms of
the strength of influence that a prior perturbation has on workers' responses.
Since workers' responses are stochastic, we define the extent of bias in 
terms of the change induced in the distribution of responses\footnote{
	This is a general divergence metric between probability 
	distributions known as the \textit{total variation distance.}
}:
\begin{equation}
	\theta = \frac{1}{2}\sum_{x \in X} \left| p_0(x) - p_1(x) \right|,
	\label{eq:theta}
\end{equation}
where $p_i(x)$ is the probability that a worker, subjected to treatment $i$,
provides response $x$, from the set of possible responses $X$.
%A significant result from a $\chi^2$ test only allows us to conclude that 
%$\theta$ is non-zero (see supplementary text).  
Determining $\theta$ from finite samples of responses, using purely 
statistical approaches, is difficult (see supplementary text),
which leads us to reformulate the
problem in terms of classification: given a worker's 
responses to test tasks, can a machine reliably infer what her prior exposure 
had been?  Intuitively, the greater the effect that prior tasks have on 
subsequent responses, the more accurate the classifier can be.  
Formally (as shown in the supplementary text), the accuracy of such a 
classifier, $\eta$, establishes a lower bound on the bias:
\begin{equation}
	\theta \geq 2\eta - 1.
	\label{l1}
\end{equation}

By measuring the accuracy of a classifier, we bound the 
bias induced by a perturbation, such as the exposure to prior tasks or 
framing.
Bias ranges from 0 (meaning the exposure 
does not affect the distribution of responses), to 1 
(or 100\%) (meaning that the exposure leads workers to provide a completely
different set of responses).

\begin{figure}
	\centering
	\includegraphics[scale=1.0]{figs/images.jpg}
	\caption{
		Examples of images used in
		initial tasks for the (\textbf{A}) \textit{food} and (\textbf{B}) 
		\textit{objects} treatments of \textit{intertask-food-objects};
		(\textbf{C}) test tasks for \textit{intertask-food-objects} and 
		\textit{frame-food-objects};
		initial tasks for the (\textbf{D}) \textit{food} and (\textbf{E}) 
		\textit{culture} treatments of \textit{intertask-food-culture};
		and (\textbf{F}) test tasks for \textit{intertask-food-culture} and 
		\textit{frame-food-culture}.
		The full set of experimental materials is shown in the 
		supplementary text.
	}

	\label{fig:task}
\end{figure}

Using a na\"ive Bayes classifier\footnote{
	We believe na\"ive Bayes is the best choice in this setting, as discussed
	in the supplementary material, and we show results when a support vector
	machine is used instead. 
	We also discuss data preprocessing and show results for alternative 
	preprocessing.
}, we found that intertask effects 
lead to a 30\% bias between workers from the \textit{food} and 
\textit{objects} treatments (Fig.~\ref{fig:theta}A).  
This represents a substantial potential to distort microtask data.

There has been considerable interest in the effects that framing can have
on microtask responses
\cite{Kinnaird2012281,chandler2013breaking,thibodeau2013natural}, so,
as a point of comparison, we describe a similar experiment 
in which we framed the purpose of the tasks.  In the experiment 
\textit{frame-food-objects}, the same test-tasks were used, but rather
than exposing workers to initial tasks, workers were either told that 
the tasks were ``Funded by the laboratory for the visual perception of Food 
and Ingredients'', or ``\ldots of Objects and Tools''.  
Framing did induce changes in the frequencies of 
word usage at significance (as determined by $\chi^2$ test; $p=0.0012$), but 
the \textit{extent} of framing-induced bias was not statistically 
distinguishable from zero ($p =0.37$) (Fig.~\ref{fig:theta}A), and was
much weaker than the bias due to intertask effects ($p=2.1\times 10^{-5}$).
\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/theta.pdf}
	\caption{
		Empirical bias, $\theta_\mathrm{NB}$, measured using a na\"ive Bayes 
		classifier, induced in image labeling tasks, due to workers' 
		exposure to framing and initial tasks.  
		(\textbf{A}) Exposing workers to initial tasks induces more severe
		bias than framing, except when echoed framing is used. 
		(\textbf{B}) As workers proceed through test tasks, 
		the bias due to initial tasks wanes, 
		but remains significant even after five tasks.  
		Standard error bars are shown.
	}
	\label{fig:theta}
\end{figure}

To assess the persistence of intertask effects, we replicated the experiment
\textit{intertask-food-objects} while permuting the test tasks, to reveal the 
dependence of intertask effects on task position.
We found that bias was strongest for the first test task (about 28\%), 
but remained significant through all five test tasks 
($\alpha=0.05$) sustaining over 12\% bias even when the initial and test 
tasks were separated by four intervening tasks (Fig. \ref{fig:theta}B).

We conducted variants of these experiments using different images, to see 
whether this trend was robust.  In the experiment 
\textit{intertask-food-culture},
workers were either assigned to a \textit{food} or \textit{culture} treatment.
The initial tasks contained either images 
depicting food (Fig.~\ref{fig:task}D), or depictions of cultural scenes 
(of dance, sport, or music) (Fig.~\ref{fig:task}E).  The test tasks, which 
were, again, identical for both treatments, depicted meals of identifiable 
cultural origin\footnote{
	Despite efforts to the contrary, initial tasks 
	in \textit{intertask-food-culture} are arguably still of
	identifiable cultural origin, but this would only tend to make our 
	results more conservative (we discuss this further in the supplementary 
	text).
}
(Fig.~\ref{fig:task}F).

Results for this experiment again showed a 
strong bias as a result of intertask effects (about 50\%) 
(Fig.~\ref{fig:theta}A).  Again, comparing intertask effects to 
framing, the experiment (\textit{frame-food-culture}) used
the same test tasks as in \textit{intertask-food-culture}, but exposed
workers to statements framing the purpose of the tasks as the recognition 
of either food or culture.  In this experiment, framing did not induce 
significant changes in word frequencies (by $\chi^2$ test, $p=0.29$) (Fig.~\ref{fig:theta}A).

It was only when framing was combined with an active reiteration step, 
that bias reached a comparable extent to that induced by intertask effects.  
In the experiment \textit{echo-food-objects},
after framing the purpose of the task (as the recognition of food
or objects), workers were asked to echo the purpose of the task
using a combo-box input.  This  ``echoed framing'' induced a bias of about 
35\% (Fig.~\ref{fig:theta}A). However, it is difficult to say whether this 
should be considered as a framing treatment \textit{per se}:
requiring the worker to reiterate the purpose signals our intent, as the 
requester, to ensure that the worker has taken note of it, possibly leading 
the worker to interpret the exchange as an instruction.  
In any case, it is remarkable that intertask effects
were on par with an explicit, actively reinforced statement of the tasks' 
purpose.

>>>>

% My packages
\usepackage{dcolumn}
\usepackage{multirow}
\newcolumntype{d}{D{.}{.}{4.0}}
\newcolumntype{s}{D{.}{.}{1.4}}
\usepackage{arydshln}
\usepackage{enumitem}

@@

\begin{abstract}
Microtask platforms are becoming commonplace tools for performing human
research, producing gold-standard data, and annotating large datasets.
These platforms connect \textit{requesters}
(researchers or companies) with large populations (crowds) of workers, who 
perform small tasks, typically taking less than five minutes each.
A topic of ongoing research concerns the design of tasks that elicit high
quality annotation.
Here we identify a feature of nearly all crowdsourcing 
workflows that profoundly impacts workers' responses.
Microtask assignments typically consist of a sequence 
of tasks sharing a common format (e.g., circle galaxies in an image). 
Using image-labeling, a canonical microtask format, we 
discover that earlier tasks have a priming effect on the worker, shifting 
the distribution of future responses by 30-50\% 
(total variational distance). 
Specifically, prior tasks influence the content that workers focus on, 
as well as the richness and specialization of responses. 
We call this phenomenon \textit{intertask effects}.
We compare intertask effects to the overt framing of the assignment, 
effected by stating the requester's research interest, 
and find that intertask effects are on par or stronger.
While intertask effects can be a source of systematic bias, 
our results suggest that, with appropriate task design, 
they might be leveraged to hone worker focus and acuity, 
helping to elicit reproducible, expert-level judgments.
Intertask effects are a crucial aspect of human computation that should be
considered in the design of any crowdsourced study.
\end{abstract}

@@

\section{Introduction}
There are many tasks that are trivial for people, but difficult to solve
programatically.  
Typical tasks include tagging and categorizing images,
%\cite{6116320,Zhai2012357}, 
coding and transcribing media,
%\cite{chandler2013breaking,Berinsky2012351,Finnerty2013},%paolacci2010running},
%judging the relevancy or quality of content,
and performing surveys for 
academic or market reserach purposes (see Table~\ref{table:task_composition} for a listing of task types seen on the Amazon's Mechanical Turk 
microtask platform).
%\cite{le2010ensuring,grady2010crowdsourcing,alonso2009can,kazai2013analysis}.
Many tasks are ill-defined, in the sense that they do not have
a clear ``correct'' response, and require high-level, qualitative judgment.
Microtask platforms are marketplaces that help fill the gap in current 
computational capabilities, by matching requesters, who need to have such 
tasks completed, with human workers.  Amazon's Mechanical Turk, and 
CrowdFlower are two popular microtask platforms.

This form of crowdsourcing embeds workers in a controlled but flexible
task infrastructure.  
To the requester, workers seem almost like input-output devices.  
This provides much of the flexibility and 
cost-savings of fully automating the work in a computer program: 
the workforce is available on demand over the Internet using automated 
scripts, without the need for interviews or contracts 
\cite{wolfson2011look,5543192}.
Work can be performed at a fraction of the cost of traditional methods for 
recruiting temporary workers or experimental subjects
\cite{Berinsky2012351}. %,ranade2012crowdsource}%,paolacci2010running}.
Many researchers consider microtask platforms 
as a new kind of \textit{human computing} architecture
\cite{5543192}. %,little2010turkit,minder2012crowdlang,kittur2011crowdforge}.

The flexibility and cost-effectiveness of microtask labour
has led to a surge in demand from industry and 
academia \cite{wolfson2011look,Berinsky2012351}.  More recently microtask 
platforms have
been assessed as a way to supplement expert human resources to increase
capacity in critical applications, such as in medical diagnostic functions,
with promising results \cite{Warby2014385}.

Naturally, researchers have investigated the factors affecting the 
reliability of microtask work, including the design of the task interface
\cite{Finnerty2013},
the design of workflows (how work is divided into tasks)
\cite{kittur2011crowdforge,Huang201077,laseckieffects},
and the framing of tasks
\cite{Kinnaird2012281,chandler2013breaking,thibodeau2013natural}.
Here we draw attention to a ubiquitous yet overlooked feature of microtask 
work: the tendency for workers to perform many similar tasks in quick 
succession.

This tendency arises from a combination of worker preferences and the 
logistics of serving small tasks.  
Workers have a preference for performing sequences of similar microtasks
\cite{Chilton20101}, probably because it reduces 
cognitive load arising from task-switching \cite{Adamczyk2004271}.
Morever, as workers complete task assignments, they must continually 
switch between working on an assignment and choosing their next 
assignment (weighing such factors as the wage paid, and effort required).  
Since microtasks are very short,
it makes sense to bundle tasks together into
larger assignments, to reduce the overhead of switching between  
working on assignments and choosing them.
This may explain why the pre-defined assignment templates available on 
microtask platforms generally bundle many tasks together by 
default\footnote{e.g. on \url{mturk.com} and \url{crowdflower.com}}.

Psychological experiments show that the exposure to stimuli immediately 
before performing a task influences performance, an effect known as 
\textit{priming} \cite{BJOP1796}.
The effect of priming appears to be stronger when the modality of the
prime is the same as the task.
Thus, the typical microtask setup creates the conditions in which
earlier tasks could induce bias in later ones, through priming.

We seek to determine if any significant priming effect occurs between
tasks, and if so, to measure and characterize it.
Certain tasks admit a well-defined notion of ``correct'' and 
``incorrect'' responses.  But many tasks involving qualitative judgment
do not.  To provide a generalized measure of bias, we seek to quantify the
extent to which intertask priming can shift the distribution of responses 
to a task.  
Measuring the \textit{extent} of the shift in response distribution,
i.e. the extent of bias, is substantially harder than
simply determining \textit{whether} a biasing effect has occured.  
Our first contribution is a statistically grounded method for doing so.

To investigate the phenomenon in a
highly generalizable setting, we sought a task that would involve 
qualitative judgment and enable relatively unconstrained responses. 
We also sought a task that would be a typical exemplar of the kinds of
tasks actually seen on microtask platforms (see 
Table~\ref{table:task_composition} for examples).  For this purpose, we 
adopted
image-labeling as a canonical micratask, in which workers provide 
descriptive labels to images using free-text inputs.  Image labeling is
a qualitative task without clearly ``correct'' or ``incorrect'' responses,
and is among the most common kinds of tasks on the major crowdsourcing
platform Amazon's Mechanical Turk.  If task-induced priming arises in our 
setup, it should be expected in virtually every microtask and 
crowdsourced workflow where workers are called upon to provide qualitative
judgment.

In our experiments, workers label a series of images, one at a time.
Depending on the treatment to which a worker is assigned, we vary the
images in the first five tasks, while keeping those in the last five tasks 
the same.  For example, in one experiment, the first five images shown to
one group of workers contained food, while those shown to the other 
contained (non-food) objects.  The last five images for both
groups contained both food and objects.

Our results show that a worker's responses are strongly influenced by 
the content of tasks performed beforehand, leading to as much as 50\%
bias (measured as total variational distance, 
see Figure~\ref{fig:l1_example} for illustrative definition).  
Using the wordnet 
knowledge base, we analyze worker's word choices to characterize the 
the nature of these effects in detail.  We find that, when workers 
label a series of images that are more similar, their responses become 
more specialized more diverse.  Prior tasks can shift the topical focus of 
worker's labels, inducing them to focus on different aspects of the images.

As a point of comparison, we perform similar experiments, in which we
alter the framing of microtask assignments, in terms of the stated purpose
of the task.  Remarkably, the effects prior microtasks themselves,
which are virtually ubiquitous, are on par with, or stronger than, 
overtly framing the purpose of an assignment.

We call the effects that earlier tasks exert on later ones 
\textit{intertask effects}.  If, as has been suggested \cite{5543192}, 
microtask platforms are to be considered as a new form of computing 
architecture, it will be necessary to reconcile the fact that the human 
computing elements exhibit \textit{hysteresis}, 
meaning that microtask workers' outputs 
depend on the \textit{history} of their inputs.
We propose a dual priming mechanism which can exaplain our
observations.  Irrespective of the underlying mechanism, this 
very strong effect might be exploited in task design to tune worker 
focus and acuity.

As our main contributions, we
\begin{itemize}[noitemsep,nolistsep]
  \item{derive a general method for measuring bias in responses;}
  \item{measure bias in microtasks due to prior tasks;}
  \item{show this effect is stronger than, or on par with framing;}
  \item{
	show that when completing a series of similar microtasks,
	worker's responses become more specialized and diverse;
  }
  \item{provide design considerations in light of our observations.}
\end{itemize}

The rest of the paper is organized as follows.  In the next section, we 
review the relevant prior work.  We then present a method for 
measuring bias due to intertask effects.  
Next we present our experimental results, followed by a discussion 
of interpretations and ramefications for microtask-based and crowdsourced 
work.  We conclude with suggestions for task design and future work.

\section{Prior work}
\subsection{Effects of microtask design on response quality}
% Accuracy, reproducibility, and a controlled task environment are 
% important
Microtask platforms are increasingly used clean and codify datasets,
administer experiments with human participants, and distribute 
surveys.  Accuracy and replicability are crucial in all of these 
applications.

\begin{table}
\centering
\begin{tabular}{l c c}
\toprule
Task Type & Count & Fraction (\%) \\
\toprule
Image transcription & 57 & 28.5 \\
Information gathering & 46 & 23.0 \\
Image tagging and classification & 36 & 18.0 \\
Copy writing and editing & 18 & 9.0 \\
Text tagging and classification & 15 & 7.5 \\
Audio/Video transcription & 12 & 6.0 \\
Survey & 9 & 4.5 \\
Unknown & 7 & 3.5 \\
\bottomrule
\end{tabular}
\caption{
	Frequency of various broad task types seen among the 200 most 
	recently posted tasks on MTurk, accessed on 12 February, 2015.
}
\label{table:task_composition}
\end{table}

% These things are definitely not assured, and people have tried to 
% figure out what to do about it
Accuracy and replicability are central to any crowdsourcing study.
There is ongoing research investigating how the desing of microtasks,
and their context, affect the quality of responses.  
One of the most basic issues with crowd work
is that workers vary in the amount of attention they pay to tasks, 
and in their understanding of instructions.
Workers who disregard or misunderstand instructions can be effectively 
screened out by including quiz-questions among the tasks, while
instruction comprehension can be improved using a training phase before
actual tasks begin
\cite{le2010ensuring,kazai2013analysis}. %paolacci2010running},
In addition, providing real-time feedback to workers encourages  
higher-quality responses, while prompting workers to review their work
according to a set of criteria increases the worker's response quality 
over time \cite{Dow20121013}.

% Designing tasks themselves
The design of the task interface also has important effects on response
quality.  A simpler user interface can improve accuracy 
\cite{Finnerty2013}, and certain input controls
are inherently less effortful and error-prone than others
\cite{cheng2015measuring}.

% Task pricing
Since microtask work is generally paid, wage is a basic parameter in
the requesters control that might influence response quality.
One might expect a higher wage to \textit{buy} better performance,
but results are ambiguous.  One study found that increasing wage
has no effect on response quality, but does increase the amount of 
responses \cite{Mason200977}.
Another study \cite{kazai2013analysis} showed that increasing
wages does increse quality, but that this effect saturates, and 
eventually increases in wage attract a larger porportion of workers who 
disregard the task instructions.

% Task framing
Workers have other motivations for completeing tasks aside from 
remuneration \cite{kazai2013analysis}.  
Researchers have attempted to appeal to
worker's motivation to do work that has a meaningful purpose:
when a task was framed as assisting with medical research,
workers are more likely to participate and completed more tasks, than
when they were not provided any context \cite{chandler2013breaking}.
Providing a meaninful context did not, however, increase the 
\textit{quality} of work.

Other studies have investigating different means of framing.  
When asked to rate various policy interventions, 
workers emphasized respectively more or less punitive approaches
depending on whether the phrase ``crime is a beast'' or 
``crime is a virus'' appeared in the problem description
\cite{thibodeau2013natural}.
When workers perform a task within the context of larger workflow, 
explaining how the task fits into the workflow, 
which amounts to a kind a framing, increased both the quantity and 
quality of responses \cite{Kinnaird2012281}.

% Optimization based on the breakdown of tasks
Breaking down of complex jobs into simpler tasks, can increasing worker
efficiency by enabling greater parallelism.
While this could disrupt the natural context provided by performing the
whole job as a single task, it turns out that, 
for complex tasks such as writing an article, 
dividing the job into subtasks, including outlining, 
information-gathering, and paragraph-synthesis, actually improves the
quality of the end-product \cite{kittur2011crowdforge}.

% effects of task interruption
Perhaps the most similar study to the work we describe here 
investigated the effects of interruptions on task performance
\cite{laseckieffects}.  There, workers 
completed a series of tasks that required them to answer questions about
an illustrated map.  When interruptions were introduced
in the form of either a time delay, or a different task involving a
different map, workers took longer to complete the task which 
followed the interruption.  However this study did not test for
effects on the quality of responses.
%\cite{kreifeldt1981interruption}

Taken together, the prior work suggests that, beyond the design
of tasks themselves, the design of the context of tasks has a major
effect on responses.
However, the potential of earlier tasks to act as primes for later ones 
has not been adequately investigated.

%\subsection{Performance during repetitive tasks}
%Studies of worker performance during repetitive work have been studied 
%in the context of long-running monitoring tasks, such as those required
%of some facility operators.
%Studies show that performance tends to decline, in terms of
%lengthening reaction time and increased rates of error, as a function of 
%time-on-task \cite{pattyn2008psychophysiological}.  In such 
%contexts, interruptions may actually serve to restore mental attention.
%Studies have shown that improved recall during tasks that were 
%interrupted.  
%
%These studies illustrate the impact of task-duration and
%repetition on performance.  However, the
%relatively low-input monitoring tasks involved in such studies are very
%different from typical microtasks, which require continuous cognitive 
%effort and input from the worker.
%
%In the microtask setting, one study investigated the effect of delays
%and interruptions introduced into a microtask setting.  This study
%found that interruptions and delays
%leading to a significant increase in the time needed to complete the
%interrupted task once resumed.
%However, the researchers did not investigate
%whether interruptions could affect the content of responses.
%
%Here we focus on how the content of earlier tasks can influence responses
%to tasks later ones.  Because qualitative judgment tasks are difficult
%to automate, but often easy for people to perform, many microtasks are of 
%this kind.  We hypothesize that such tasks are highly susceptible to 
%priming effects, and that in the repetitive pattern of microtask work
%(and most crowdsourced work), earlier tasks may act as primes to later 
%ones, and thereby introduce singificant amounts of bias.

\subsection{Priming}
Priming is a psychological phenomenon whereby previous exposure to a 
stimulus leads to faster or more accurate responses to similar stimuli 
\cite{Ghuman17062008},
or to lower response thresholds \cite{BJOP1826}.  
Priming occurs by means of pre-activation, which can influence
stimulus encodings (during perception) \cite{BJOP1826}, 
the top-down application of object-knowledge \cite{Ghuman17062008},
and memory access \cite{beller1971priming}.
Priming is more 
effective when the prime and the target task involve the kind of 
activity.  For example, exposing participants to an image helps them 
subsequently recognize that image when shown very breifly in a 
tachistoscope; however, being presented with a related \textit{word}, 
and reading it aloud, does not \cite{BJOP1796}.

As we have mentioned, workers prefer to perform series of similar tasks
series of similar tasks in quick succession.  Moreover, the many 
microtasks involve tasks based on visual and auditory perception 
similar to the kinds used in priming studies.
Thus, in a typical microtask assignment, there are many opportunities for 
priming, the effects of which should be explored.

\subsection{Measuring the extent of priming}
If a prior task acts as a prime for a later one, it will 
result in a shift in the responses workers provide for the later task.
A variety of tests exist to determine \textit{whether} two samples, 
i.e. sets of responses, come from different distributions, 
most notably the $\chi^2$ test.  
This will suffice to determine whether there are any grounds for conern 
about priming effects in the first place.  However, such tests
do not tell us \textit{how much} two distributions differ, and hence
how severe the priming effects are.

Determining the severity of priming effects amounts to measuring a 
distance, or \textit{divergence}, between response distributions, by
sampling responses from workers exposed to different priming conditions.
We will present a method for bounding a kind of divergence known as 
\textit{total variational distance}\footnote{
  Other formulation such as the Kulback-Leibler and 
  Jenson-Shannon divergence can be calculated from the total variational
  distance
}, which we will henceforth denote by $\theta$.
The value of $\theta$ between two distributions $P$ and $Q$ 
is the fraction under the graph of $P$ that does not overlap with that
under $Q$ (or vice-versa; see Figure~\ref{fig:l1_example}).  
Formally:
\begin{align}
	\theta = \frac{1}{2}\sum_{x \in X} \left| p(x) - q(x) \right|,
	\label{eq:theta}
\end{align}
Where $p(x)$ and $q(x)$ are the respective probabilities 
that a worker submits response $x$, from a set of possible responses 
$\mathcal{X}$, under distributions $P$ and $Q$.
The divergence is constrained to 
$0 \leq \theta \leq 1$ (or from 0 to 100\%).
When $\theta = 100\%$, the distributions do not overlap at all.  
When $\theta = 0\%$, the distributions are identical.
For reference, the distributions plotted in 
Figure~\ref{fig:l1_example}A and B have $\theta=30\%$ and $\theta=50\%$ 
respectively.

\begin{figure}
	\centering
	\includegraphics[scale=1.0]{figs/normal_example.pdf}
	\caption{
	  	Normal distributions exhibiting a total variational distance
		of A) 50\% and B) 30\%.  The total variational distance is 
		the non-overlapping portion of the distributions, e.g. in
		A) it is given by 
		$\theta = a_1 / (a_1 + a_2) = a_2 / (a_2 + a_3)$.
	}
	\label{fig:l1_example}
\end{figure}

Empirically measuring the divergence between distributions
is a matter of ongoing
research \cite{val-thesis,batu2013testing,chan2014optimal}.
In the ``na\"ive'' approach one first uses the samples to reconstruct 
empirical
distributions, by calculating their maximum likelihood parameters.
Then, one substitutes the empirical values of $p(x)$ and $q(x)$ into 
Equation~\ref{eq:theta} \cite{batu2013testing}.  
For example, one might directly estimate $\hat{p}(x) = N_x/N$, 
where $N_x$ is the number of times response $x$ is observed among $N$ 
total responses.
